---
title: "Employee Data"
author: "Olaiya Martins"
date: "2024-01-11"
output: html_document
---

### Library Installed

```{r}
library(dplyr)
library(tidyverse)
library(readxl)
```

## data loaded

```{r}
employee_data <- read_excel("~/Documents/employee_data.xlsx")
menu_employee <- read_excel("~/Documents/menu_employee.xlsx")
members_employee <- read_excel("~/Documents/members_employee.xlsx")
```

### View the loaded data

```{r}
print(employee_data)
print(menu_employee)
print(members_employee)
```

### change the Order date to Visit Date

```{r}
employee_data2<-employee_data %>%
  rename(visit_date=order_date)
head(employee_data2)
```

## **Analysing the Data**

### 1. how many days has each customer visited the restaurant:

```{r}
days_visited <- employee_data2 %>% group_by(customer_id) %>% summarise(n_distinct(visit_date))
head(days_visited)
```

### 2. what is the total Amount each customer spent at the restaurant?

```{r}
merged_data <- merge(employee_data2, menu_employee, by = "product_id")
print(merged_data)

total_amount_spent <- merged_data %>% group_by(customer_id) %>% summarise(total_spent = sum(price)) # Print the result print(total_amount_spent)
print(total_amount_spent)
```

### 3. what was the first item from the menu purchased by each customer?

```{r}
first_purchases <- merged_data %>%
  arrange(merged_data$visit_date,desc(product_name)) %>%
  group_by(customer_id) %>%
  slice(1) %>%
  select(customer_id, product_name)
print(first_purchases)
```

### 4. what is the most purchased item on the menu and how many times was it purchased by all customers

```{r}
most_purchased_item <- merged_data %>%
  group_by(product_name) %>%
  summarise(purchase_count = n()) %>%
  arrange(desc(purchase_count)) %>%
  slice(1)
print(most_purchased_item)
```

### 5. Which item was the most popular for each customer?

```{r}
most_popular_item <- merged_data %>%
  group_by(customer_id, product_name) %>%
  summarise(purchase_count = n()) %>%
  top_n(1, purchase_count) %>%
  arrange(customer_id)
print(most_popular_item)
```

### 6. Which item was purchased first by the customer after they became a member?

```{r}
members_employee$customer_id <- toupper(members_employee$customer_id)

first_purchase_after_membership <- merged_data %>%
  inner_join(members_employee, by = "customer_id") %>%
  filter(visit_date > join_date) %>%
  arrange(customer_id) %>%
  group_by(customer_id) %>%
  slice(1)
print(first_purchase_after_membership)
```

### 7. Which item was purchased just before the customer became a member?

```{r}
purchase_before_membership <- merged_data %>%
  inner_join(members_employee, by = "customer_id") %>%
  filter(visit_date < join_date) %>%
  arrange(customer_id, visit_date) %>%
  group_by(customer_id) %>%
  slice_tail(n = 1)

print(purchase_before_membership) ###### A has two product name, but only one came, there should be Sushi for A also.(stuck on this)
```

### 8. what is the total items and amount spent for each member before they became a member?

```{r}
totals_before_membership<- merged_data %>%
  inner_join(members_employee, by = "customer_id") %>%
  filter(visit_date < join_date) %>%
  group_by(customer_id) %>%
  summarise(total_items = n(), Total_amount_spent = sum(price))
print(totals_before_membership)
```

### 9. if each \$1 spent equates to 10 points and sushi has a 2x points multiplier—how many points would each customer have?

```{r}
##### define the points multiplier for shushi
sushi_multiplier<- 2

##### calculate the points for each customer
points_per_dollar <- 10
points_per_dollar_sushi <- points_per_dollar * sushi_multiplier

customer_points <- merged_data %>%
  mutate(points = ifelse(product_name== "sushi", price * points_per_dollar_sushi, price * points_per_dollar)) %>%
  group_by(customer_id) %>%
  summarise(total_points = sum(points))
print(customer_points)
```

### 10. in the first week after a customer joins the program (including their join date) they earn 2x points on all items, not just sushi-how many points do customer A and B have at the end of January?

```{r}
### define the points multiplier
base_multiplier <- 1
sushi_multiplier
first_week_multiplier <- 2
january_multiplier <-2

#### points for each customer at the end of january: to work on better
customer_points <- merged_data %>%
  inner_join(members_employee, by = "customer_id") %>%
  mutate(week_after_join = visit_date - join_date < 7) %>%
  mutate(points = ifelse(week_after_join, price * first_week_multiplier, price * base_multiplier)) %>%
  group_by(customer_id) %>%
  summarise(total_points = sum(points))
print(customer_points) ###### stuck on this too and unsure of the answer
```

### 11. Recreate table with column: customer_id, order_date, product_name, price,member(Yes/No)

```{r}
joined_data <- merge(merged_data, members_employee, by = "customer_id", all.x = TRUE) %>%
   mutate(membership = ifelse(join_date <= visit_date,"Yes", "No")) %>%
arrange(customer_id, visit_date, product_name)
print(joined_data)
```

### 12. Recreate table with column: customer_id, order_date, product_name, price,member(Y/N), ranking(null/number)

```{r}
ranking <- merged_data %>%
  left_join(members_employee, by = "customer_id") %>%
  mutate(membership = ifelse(visit_date>= join_date, "Yes", "No")) %>%
  arrange(customer_id, visit_date,product_name) %>%
  mutate(ranking = ifelse(membership== "Yes", rank(visit_date),NA))
print(ranking)
```

### **Insights From the Analyzed Data**:

#### • The most popular menu is ramen. 
#### • Customers A and B are loyal customers and they become members after ordering sushi. Offer sushi to customer C to become members
